---
title: "Mice gut microbiome data analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mice gut microbiome data analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 5)
```

This vignette reproduces the real-data analysis of paper §5. The data come from
the gnotobiotic mouse experiment of Patnode et al. (2019): a 15-member human gut
bacterial consortium under a HiSF diet supplemented with different fibres, with
two categorical covariates — **diet** (HiSF / citrus pectin CPT / pea fibre PEF)
and the **presence/absence of WH2** (*B. cellulosilyticus*, OTU 2).

```{r setup}
library(BCAIA)
data(mice)
dim(mice$Y)
mice$species
```

`mice$Xmean` and `mice$Xcov` are the design matrices used in the paper
(`Xcov` includes an intercept column).

## Fitting the model

The paper uses `K = 8` and `niter = 160000` (~22 min on an M1 Mac). Here we run
a short chain for illustration.

```{r fit, eval = FALSE}
fit <- bcaia(mice$Y, mice$Xmean, mice$Xcov, K = 8,
             niter = 160000, burnin = 80000, thin = 10, seed = 1)
```

```{r fit-short, echo = FALSE}
fit <- bcaia(mice$Y, mice$Xmean, mice$Xcov, K = 8,
             niter = 1500, burnin = 750, thin = 5, seed = 1, verbose = FALSE)
fit
```

## Condition-specific interactions

Each row of `Xcov` corresponds to a sample; the six experimental conditions are
the unique rows. We estimate the posterior correlation matrix for each.

```{r conditions}
conds <- unique(mice$Xcov)
rho_list <- lapply(seq_len(nrow(conds)), function(i) posterior_cor(fit, conds[i, ]))
length(rho_list)
```

## Interaction between a pair of species across conditions

For example, OTUs 9 and 13 (*Collinsella aerofaciens* and *Ruminococcaceae*),
reported as negatively associated across all conditions in the paper:

```{r pair}
xlist <- setNames(lapply(seq_len(nrow(conds)), function(i) conds[i, ]),
                  paste0("cond", seq_len(nrow(conds))))
posterior_cor_pair(fit, j = 9, k = 13, xlist = xlist)
```

## Differential abundance

Posterior draws of the mean-regression coefficients `beta` quantify how
abundance shifts with diet and WH2 status (paper Fig. 5). For example, the
posterior of the difference between two `Xmean` columns:

```{r beta}
beta <- fit$samples$beta          # J x Pmean x nsamp
diff <- beta[, 1, ] - beta[, 2, ] # contrast of columns 1 and 2 of Xmean
data.frame(OTU = mice$species,
           median = round(apply(diff, 1, median), 2),
           lower  = round(apply(diff, 1, quantile, 0.025), 2),
           upper  = round(apply(diff, 1, quantile, 0.975), 2))
```

With the full chain these recover the significant diet and WH2 effects reported
in the paper.
