---
title: "Simulation 2: repeated samples and inter-subject variability"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation 2: repeated samples and inter-subject variability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
```

This vignette illustrates the **subject-indexed** model (paper §4.2), used when
there are repeated samples per subject and large inter-subject variability. The
baseline abundance is subject-specific ($\alpha_{s,j}$), selected automatically
by supplying a `subject` vector to `bcaia()`.

The paper's Simulation 2 uses $S = 25$ subjects, $N = 50$ samples, $J = 100$
OTUs, a continuous subject covariate and a binary sample covariate, with a
baseline covariance plus a covariate-varying part. Here we use a smaller design
so the vignette runs quickly.

```{r data}
library(BCAIA)
set.seed(3)
s <- 15; n <- 2 * s; J <- 40
subject <- rep(seq_len(s), 2)
xd <- rep(c(0, 1), each = s)     # binary sample covariate (condition)
xc <- rep(rnorm(s), 2)           # continuous subject covariate
Xcov  <- cbind(1, xd, xc)        # covariance design (with intercept)
Xmean <- cbind(xd, xc)           # mean design (no intercept)

## true covariate-varying covariance + subject baseline abundance
Q <- matrix(0, J, 2); Q[sample(J, J / 2), ] <- rnorm(J)
Fm <- matrix(rnorm(6), 2, 3)
alpha_s <- matrix(rnorm(s * J, 3, 1), s, J)
Y <- matrix(0, n, J)
for (i in seq_len(n)) {
  Lam <- Q * matrix(as.numeric(Fm %*% Xcov[i, ]), J, 2, byrow = TRUE)
  Sig <- tcrossprod(Lam) + diag(0.25, J)
  L <- chol(Sig + diag(1e-8, J))
  Y[i, ] <- floor(exp(alpha_s[subject[i], ] + as.numeric(crossprod(L, rnorm(J)))))
}
round(mean(Y == 0), 3)  # zero rate
```

## Fitting the subject-indexed model

Supplying `subject` switches on the $\alpha_{s,j}$ variant. **The paper uses
`K = 7` and `niter = 160000`**; a short chain is shown here.

```{r fit, eval = FALSE}
fit <- bcaia(Y, Xmean, Xcov, subject = subject, K = 7,
             niter = 160000, burnin = 80000, thin = 10, seed = 1)
```

```{r fit-short, echo = FALSE}
fit <- bcaia(Y, Xmean, Xcov, subject = subject, K = 6,
             niter = 1000, burnin = 500, thin = 5, seed = 1, verbose = FALSE)
fit
```

## Covariate-varying covariance for a pair of features

Because `Xcov` includes a continuous covariate, the covariance varies smoothly
with it. We evaluate the posterior covariance between two features across a grid
of the continuous covariate at each level of the binary covariate.

```{r curve, fig.width = 6, fig.height = 4}
grid <- seq(-2, 2, length.out = 25)
pair <- c(1, 2)
curve_xd <- function(d) sapply(grid, function(g)
  posterior_Sigma(fit, c(1, d, g))[pair[1], pair[2]])
plot(grid, curve_xd(0), type = "l", col = "red", ylim = range(curve_xd(0), curve_xd(1)),
     xlab = "continuous covariate", ylab = expression(Sigma[jk](x)))
lines(grid, curve_xd(1), col = "blue")
legend("topright", c("xd = 0", "xd = 1"), col = c("red", "blue"), lty = 1, bty = "n")
```

With the full chain, the model recovers the baseline covariance, identifies
inactive features, and tracks the covariate-varying interactions (paper Fig. 2).
