---
title: "Simulation 1: covariate-varying interactions with a categorical design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation 1: covariate-varying interactions with a categorical design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
```

This vignette reproduces the design of **Simulation 1** (paper §4.1): a
categorical design with a binary and a ternary covariate giving six
experimental conditions, `J = 15` features and `N = 30` samples. The true
covariance is rank-2 and varies with the covariates.

```{r setup}
library(BCAIA)
set.seed(6)
sim <- simulate_bcaia(n = 30, J = 15, K_true = 2, seed = 6)
dim(sim$Y)
round(mean(sim$Y == 0), 3)   # zero rate
```

## Choosing K

`K` is set to a value large enough to capture the covariance; redundant factors
are shrunk out by the Dirichlet–Horseshoe prior. A data-driven starting point is
the clr-PCA scree rule:

```{r choose-k}
choose_K(sim$Y)
```

The paper uses `K = 5` here (a small over-specification is harmless).

## Fitting the model

For illustration we run a short chain. **The paper uses `niter = 160000`**;
increase `niter` for real analyses (best run on a server).

```{r fit, eval = FALSE}
fit <- bcaia(sim$Y, sim$Xmean, sim$Xcov, K = 5,
             niter = 160000, burnin = 80000, thin = 10, seed = 1)
```

```{r fit-short, echo = FALSE}
fit <- bcaia(sim$Y, sim$Xmean, sim$Xcov, K = 5,
             niter = 1500, burnin = 750, thin = 5, seed = 1, verbose = FALSE)
fit
```

## Posterior covariance vs. truth

We compare the posterior median covariance to the truth at two conditions.

```{r sigma}
x1  <- sim$Xcov[1, ]    # condition 1
x26 <- sim$Xcov[26, ]   # condition 6
S1  <- posterior_Sigma(fit, x1)
S26 <- posterior_Sigma(fit, x26)

ut <- upper.tri(S1, diag = TRUE)
c(cond1 = cor(S1[ut],  sim$truth$Sigma[, , 1][ut]),
  cond6 = cor(S26[ut], sim$truth$Sigma[, , 26][ut]))
```

```{r heatmap, fig.width = 7, fig.height = 3.5}
op <- par(mfrow = c(1, 2), mar = c(2, 2, 2, 1))
image(S1,  main = "Sigma-hat (condition 1)",  axes = FALSE)
image(S26, main = "Sigma-hat (condition 6)",  axes = FALSE)
par(op)
```

With the full `niter = 160000` chain, the posterior estimates closely match the
true covariate-varying covariance across all six conditions (paper Fig. 1).
